# Chapter 6. Analysis of longitudinal data

## Loading the datasets
```{r, message = FALSE}
library(ggplot2); library(dplyr); library(tidyr)
RATSL <- read.csv(file = "C:/Users/vinnu/Documents/IODS-project/data/RATSL.csv")
RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)
str(RATSL)
BPRSL <- read.csv(file = "C:/Users/vinnu/Documents/IODS-project/data/BPRSL.csv")
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)
str(BPRSL)
```

In RATS we have three groups on differnet diets.


## Analyses on RATS 

### Graphs
``` {r}
# Graphs of RATS
ggplot(RATSL, aes(x = Time, y = weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$weight), max(RATSL$weight))) +
  ggtitle("Individual response profiles by group for the RATSL data")
```

### Standardized graphs
``` {r}
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stweigth = (weight - mean(weight))/sd(weight) ) %>%
  ungroup()
glimpse(RATSL)
ggplot(RATSL, aes(x = Time, y = stweigth, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:16, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(name = "standardized weigth") +
  ggtitle("Individual response profiles by group for RATSL data after standardization")
```

### Summary graph
``` {r}
# Number of time, baseline (week 0) included
n <- RATSL$Time %>% unique() %>% length()
# Summary data with mean and standard error of bprs by treatment and week 
RATSS <- RATSL %>%
  group_by(Group, Time) %>%
  summarise(mean = mean(weight), se = sd(weight)/sqrt(n) ) %>%
  ungroup()
# Glimpse the data
glimpse(RATSS)
# Plot the mean profiles
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.4) +
  theme(legend.position = c(0.8,0.4)) +
  scale_y_continuous(name = "mean(weigth) +/- se(weigth)") +
  ggtitle("Mean response profiles for the three groups in the RATS data.")
```
  
The mean profiles plot shows that the error bars do not overlap.



## Analyses on BPRS data

### Plotting
```{r}
# plotting again 
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```
  
From the plot we see, that between the two treatments there seem to no clear difference. With both bprs measure decreases with time, so treatments seem to work.

### Linear regression model
```{r}
# let's fit a linear regression
BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)
# print out a summary of the model
summary(BPRS_reg)
BPRS_reg
```
Treatment2 does not seem to differ from treatmet1.